A Memory-Bounded, Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-02-19 01-intro.tex
Pequenos ajustes no fim da Introdução, Figura ilegível, e Título.

In this work, we present a formal semantics for a subset of \CEU that focus on
its peculiarities in comparison to other synchronous languages.
\begin{itemize}
    \item qual a abordagem / operational semantics / dois passos
    \item quais os resultados / provas
    \item quais os desafios e limitações
    \item \gl{TODO}
\end{itemize}

\fs{Descrever seções.}

>>>>

In this work, we present a formal small-step structural operational semantics
for \CEU and proofs that it enforces memory-bounded, deterministic, and
terminating reactions to the environment, i.e., that for a given arbitrary
timeline of input events, multiple executions of the same program always react
in bounded time and arrive at the same final finite memory state.

% TODO: \fs{Descrever seções.}

### 2018-02-19 02-ceu.tex
Revisão da semântica.
Pequenos ajustes no fim da Introdução, Figura ilegível, e Título.

\begin{figure}[b]
\centering
\includegraphics[width=\columnwidth]{tick}
\caption{The discrete notion of time in \CEU.}
\label{fig.ticks}
\end{figure}

The sequential processing of external input events induces a discrete notion of
time in \CEU, as illustrated in Figure~\ref{fig.ticks}.
%
The continuous timeline shows an absolute reference clock with ``physical
timestamps'' for the event occurrences (e.g., event~\code{C} occurs at
$17ms521us$).
%
The discrete timeline shows how the same occurring events fit in the logical
notion of time of \CEU.
%
The boot reaction \code{boot-0} happens before any input, at program startup.
%
Event~\code{A} ``physically'' occurs during \code{boot-0} but, because time
is discrete, its corresponding reaction only executes afterwards, at logical
instant~\code{A-1}.
%
Similarly, event~\code{B} occurs during~\code{A-1} and its reaction is
postponed to execute at~\code{B-2}.
%
Event~\code{C} also occurs during~\code{A-1} but its reaction must also wait
for~\code{B-2} to execute and so it is postponed to execute at~\code{C-3}.
%
Event~\code{D} occurs during an idle period and can start immediately at
\code{D-4}.
%
Finally, two instances of event~\code{E} occur during~\code{D-4}; they are
handled in the subsequent reactions~\code{E-5} and~\code{E-6}.

@@

\begin{lstlisting}[
  numbers=left,
  basicstyle=\ttfamily\footnotesize,
  float=h,
  caption={A \CEU program with a ``subroutine''.},
  label={lst.sub},
]
event int* inc;         // declares subroutine "inc"
par/or do
    var int* p;
    every p in inc do   // implements "inc" through an event iterator
        *p = *p + 1;
    end
with
    var int v = 1;
    emit inc(&v);       // calls "inc"
    emit inc(&v);       // calls "inc"
    _assert(v==3);      // asserts result after the two returns
end
\end{lstlisting}
\vskip-\baselineskip

>>>>

\begin{figure}[b]
\centering
\includegraphics[width=\columnwidth]{tick_min}
\caption{The discrete notion of time in \CEU.}
\label{fig.ticks}
\end{figure}

The sequential processing of external input events induces a discrete notion of
time in \CEU, as illustrated in Figure~\ref{fig.ticks}.
%
The continuous timeline shows an absolute reference clock with ``physical
timestamps'' for the event occurrences (e.g., event~\code{C} occurs at
$17ms521us$).
%
The discrete timeline shows how the same occurring events fit in the logical
notion of time of \CEU.
%
The boot reaction \code{boot-0} happens before any input, at program startup.
%
Event~\code{A} ``physically'' occurs during \code{boot-0} but, because time
is discrete, its corresponding reaction only executes afterwards, at logical
instant~\code{A-1}.
%
Similarly, event~\code{B} occurs during~\code{A-1} and its reaction is
postponed to execute at~\code{B-2}.
%
Event~\code{C} also occurs during~\code{A-1} but its reaction must also wait
for~\code{B-2} to execute and so it is postponed to execute at~\code{C-3}.
%
Finally, event~\code{D} occurs during an idle period and can start immediately
at \code{D-4}.
%
%Finally, two instances of event~\code{E} occur during~\code{D-4}; they are
%handled in the subsequent reactions~\code{E-5} and~\code{E-6}.

@@

\begin{lstlisting}[
  numbers=left,
  basicstyle=\ttfamily\footnotesize,
  float=h,
  caption={A \CEU program with a ``subroutine''.},
  label={lst.sub},
]
event int* inc;         // declares subroutine "inc"
par/or do
    var int* p;
    every p in inc do   // implements "inc" through an event iterator
        *p = *p + 1;
    end
with
    var int v = 1;
    emit inc(&v);       // calls "inc"
    emit inc(&v);       // calls "inc"
    _assert(v==3);      // asserts result after the two returns
end
\end{lstlisting}
%\vskip-\baselineskip

### 2018-02-19 index.tex
Adição do apêndice.
Adiciona ao apêndice prova (incompl.) da terminação NST
Pequenos ajustes no fim da Introdução, Figura ilegível, e Título.

\documentclass[sigplan, anonymous, review]{acmart}

@@

\title[A Memory-Bounded, Deterministic, and Terminating Semantics for \CEU]
{A Deterministic and Terminating Semantics
  for the Synchronous Programming Language~\CEU}

@@

\appendix
%\include{proofs}

>>>>

\documentclass[sigplan,anonymous,review,protrusion=true,expansion]{acmart}

@@

\title{A Memory-Bounded, Deterministic and Terminating Semantics for the
       Synchronous Programming Language~\CEU}

@@

\input{06-appendix}
